Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-03 WSSSPE1_report/WSSSPE1_summary_report.tex
Combined a few points to make sections 4.2 less fragmented.

He noted that in places, development in science has improved thanks to
open source and Github, but for the most part remains arcane. He argues
that there's a lot we can learn from the app-store model. He
mentioned BioJava\footnote{\url{http://biojava.org/wiki/Main_Page}} and
OpenScienceDataCloud \footnote{\url{https://www.opensciencedatacloud.org/}} as
distribution channels.

@@

He has problems with staff retention with respect to private organizations which reward 
those combining research and software expertise (the ``Google Bus''), however,
he's a strong fan of private sector sustainability sources and PPP. He
notes that making a successful business from scientific software alone
is very rare: founders overvalue while customers undervalue. He noted
that to last, an open source project needs a minimal funding requirement
even with a vibrant community -- good will only goes so far if you're
being paid to do something else.

@@

He had problems with selling research software: his technology transfer
office wanted wanted huge IP reachthrough, whereby they would get a
share of profits from drugs developed by pharma companies who use the
software. He was aware this was unrealistic but the technology transfer
office insisted for a
while. He wants to push a much more one-click approach to purchasing
university-written software.

@@

He asked, of this practice, ``why didn't we know better?'' and argued
this was because we're taught to focus on immediate research outcomes
and not on continuously improving and building on tools for research. He
then asked, later in our working lives, when we do know better, why
don't we act any different, and argued this was due to incentives and their lack:
only the immediate products of research, not the software, are valued.

He referenced Victoria Stodden's talk at OKCon\footnote{\url{http://okcon.org/main-speakers/victoria-stodden/}},\footnote{\url{http://www.stanford.edu/~vcs/talks/OKcon2013-Sept172013-STODDEN.pdf}}
which he said argued these points well.

He cited the EMCEE\footnote{\url{https://github.com/dfm/emcee}} Markov Chain
Monte Carlo code by Dan Forman Mackey\footnote{\url{http://dan.iel.fm/}} as a
good example of a scientific project using GitHub. Amongst other
examples of good scientific projects on GitHub he used an example from Joe Zuntz\footnote{\url{http://www.jb.man.ac.uk/~zuntz/Joe_Zuntz/Welcome.html}}.

C. Titus Brown\footnote{\url{http://ged.msu.edu/}} argues that with regard to
reusable software, ``we should just start doing it''. Arfon notes that
documentation should be ``treated as a first class entity.''

He noted that some software fields have excellent cultures of module
reuse, such as rubygems\footnote{\url{http://rubygems.org/}},
PyPI\footnote{\url{https://pypi.python.org/pypi}}, and CPAN\footnote{\url{http://www.cpan.org/}}, where there is effectively
ultra-low-friction collaboration through these repositories. This has
not happened in highly numerical, compiled language scientific software.

He argues that the GitHub Pull Request code review mechanism facilitates
such collaboration, by allowing one to code first, and seek review and a
merge back into the trunk later.

Fernando Perez says that ``Open Source
is reproducible by necessity'' - since reproducibility is a prerequisite
for remote collaboration.

Arfon therefore defines GitHub as part of the next stage in development
of the web, following on from the ``social web'' of Facebook et.al.:
``The Collaborative Web.''

>>>>

He noted that in places, development in science has improved thanks to
open source and Github, but for the most part remains arcane. He argues
that there is a lot we can learn from the app-store model. He
mentioned BioJava\footnote{\url{http://biojava.org/wiki/Main_Page}} and
OpenScienceDataCloud \footnote{\url{https://www.opensciencedatacloud.org/}} as
distribution channels.

@@

He experienced problems with staff retention with respect to private organizations which reward 
those combining research and software expertise (the ``Google Bus''). However,
he is a strong fan of private sector sustainability sources and PPP. He
noted that making a successful business from scientific software alone
is very rare: founders overvalue while customers undervalue. He noted
that to last, an open source project needs a minimal funding requirement
even with a vibrant community -- good will only goes so far if you are
being paid to do something else.

@@

He had problems with selling research software: the technology transfer
office in his university wanted huge IP reachthrough, whereby they would get a
share of profits from drugs developed by pharma companies who use the
software. He was aware this was unrealistic but the technology transfer
office insisted for a
while. He wants to push a one-click approach to purchasing
university-written software.

@@

He asked, of this practice, ``why didn't we know better?'' and argued
this was because we were taught to focus on immediate research outcomes
and not on continuously improving and building on tools for research. He
then asked, later in our working lives, when we do know better, why
don't we act any different, and argued this was due to incentives and their lack:
only the immediate products of research, not the software, are valued.
He referenced Victoria Stodden's talk at OKCon\footnote{\url{http://okcon.org/main-speakers/victoria-stodden/}},\footnote{\url{http://www.stanford.edu/~vcs/talks/OKcon2013-Sept172013-STODDEN.pdf}}
which he said argued these points well.

He cited the EMCEE\footnote{\url{https://github.com/dfm/emcee}} Markov Chain
Monte Carlo code by Dan Forman Mackey\footnote{\url{http://dan.iel.fm/}} and Joe Zuntz\footnote{\url{http://www.jb.man.ac.uk/~zuntz/Joe_Zuntz/Welcome.html}} as 
good examples of  scientific projects using GitHub.

C. Titus Brown\footnote{\url{http://ged.msu.edu/}} argues that with regard to
reusable software, ``we should just start doing it.'' Arfon noted that
documentation should be ``treated as a first class entity.''
He noted that some software fields have excellent cultures of module
reuse, such as rubygems\footnote{\url{http://rubygems.org/}},
PyPI\footnote{\url{https://pypi.python.org/pypi}}, and CPAN\footnote{\url{http://www.cpan.org/}}, where there is effectively
ultra-low-friction collaboration through these repositories. This has
not happened in highly numerical, compiled language scientific software.

He argued that the GitHub Pull Request code review mechanism facilitates
such collaboration, by allowing one to code first, and seek review and a
merge back into the trunk later.

Fernando Perez says that ``Open Source
is reproducible by necessity'' - since reproducibility is a prerequisite
for remote collaboration.
Arfon therefore defined GitHub as part of the next stage in development
of the web, following on from the ``social web'' of Facebook et.al.:
``The Collaborative Web.''
